---
title: "Figma Page views (TimeSeries) with R"
author: "Salami Famous (salamifamous@gmail.com)"
date: "2023-01-15"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## loading useful libraries
```{r include = F} 
# code block checking working directory and loading typical libraries

library(ggplot2) # for graph plots
library(stringr) #for string manipulations
library(dplyr)   # needed for data manipulations
library(prophet) # The forecast library
library(lubridate)
library(textshape)
library(forecast)
library(melt)
library(reshape)
library(tidyverse)
theme_set(theme_classic())
```

```{r}
#clear environment
rm(list=ls())
```

```{r}
#set working directory
setwd(getwd())
```

# Task (Page views prediction)

```{r}
#Reading the data file
figmadata <- read.csv("data/figma-views-20220101-20220823.csv", stringsAsFactors = T)

```

## Transforming the 'Date' column to standard date format
```{r}
#analyse the data

figmadata$Date = as.Date(figmadata$Date,format="%m/%d/%Y")
str(figmadata)

```
```{r}
# Parse date column, set the date column as index

figmadata <- mutate (
figmadata,
date = ymd_hms(figmadata$Date) # parse date column using lubridate ymd_hms function
)

```
```{r}
figmadata <- mutate (
figmadata,
ds = figmadata$Date,  # Create new ds column from date using mutate
y = figmadata$Pageviews   # Create new y column from value using mutate
)
```
## Analysis of the historical data using simple plot
```{r}
#ggplot(data = figmadata, aes(x = Date, y = Pageviews)) + geom_boxplot()

ggplot(data = figmadata, aes(x = Date, y = Pageviews)) + geom_line(color = "#00AFBB", size = 1, na.rm = TRUE) + theme(axis.text.x = element_text(angle = 90))


# Allow Default X Axis Labels
# Allow Default X Axis Labels
ggplot(figmadata, aes(x=Date, group = 1)) + 
  geom_line(aes(y=Pageviews, group = 1), na.rm = TRUE) + 
  labs(title="Time Series Chart", 
       subtitle="Daily page views from Figma Dataset", 
       caption="Source: Figma", 
       y="Page views")

```

## We are ready to fit a model to the historical data
```{r, echo=FALSE}
#figmadata_clean = figmadata[!is.na(figmadata)]

tmodel <- prophet(figmadata,yearly.seasonality = "auto", daily.seasonality = "auto", seasonality.prior.scale=0.9)

```
```{r}
#date range from now till middle of next year

date_from<-as.Date("2023-01-15")
date_to<-as.Date("2024-06-30")

daterange = seq(from = date_from, to = date_to, by = 'day')
predDuration = as.numeric(length(daterange)) #533
originalDuration = as.numeric(count(figmadata))
totDuration = as.numeric((originalDuration + predDuration - 94))

```

## Prediction fo rfuture dates
```{r}
future <- make_future_dataframe(tmodel, periods = totDuration)
```

## Model fitting
```{r}
forecast <- predict(tmodel, future)
```

```{r}

pred_forecast = tail(forecast, predDuration)
pred_forecastt = tail(forecast, predDuration)

tail(pred_forecast)

```


## Combined Graph of historical and predicted data. The historical data is shown in dots with the predicted views represented in line graph

```{r}
plot(tmodel, pred_forecast)
```




## Graph showing trend of Forecast (predicted) views, see below
```{r}
prophet_plot_components(tmodel, pred_forecast)
```

Unlike the Python version where I set DAILY SENSATIONALITY of the forecast to TRUE, here, I set it to default, that is, the forecast library takes decision on its own. This is why we have just TWO graphs above as against three.

I need to plot this graph to show the trend or progression of the page views monthly and weekly
We can see how the trend (page views) behaves, rising and falling depending on the vi
Graph 1 - Shows the predicted rise in views from the beginning of 2023 and peaks to 2
Graph 2 - Mondays to Thursdays show higher predicted page views while it declines on 

# '

## Dotplot showing the progression trend for predicted page views, see below
```{r}
ggplot(data = pred_forecast, aes(x = ds, y = yhat, col = yhat)) + geom_point()

```